Enhancing Forecast Reconciliation: A Study of Alternative Covariance Estimators

### 2025-08-01 literature/paper_draft.qmd
refine 2nd sim setting

In this section we explore the effects of "sparse" and "dense" correlation matrix in our data generating process for series at the bottom level. We consider a large structure, having 2 groups of 50 bottom series. The cross-series within-group correlations are strong, and the between-group correlations will be weaker for the "dense" case, and zero for the "sparse" case. The dense and sparse correlation matrices for the DGP of bottom series are illustrated in @fig-Sigma-50x2. Note that these 2 cases have the same VAR(1) structure, only the correlation matrix is different.

However, rather than aggregating series within group to form 2 level-1 series as we did earlier, we combine them in groups of size 10 to obtain 10 level-1 series. These 10 series will be splitted into 3 groups, having 4, 3, and 4 series respectively, and are aggregated into 3 level-2 series. Finally, they are added to obtain the top level. The intention behind this unusual aggregation path is to create a more complex structure for higher levels, in which the level-2 series will correlate with the bottom series in the first or second group, or both.

Results are presented in @fig-sim-results-dsp, with the graphs in the first column being the "dense" case and the ones in second column being "sparse" case. We consider two time series length--54 and 304 observations with the last 4 being the test set--presenting at the first and second row respectively. As shown in the plots, MinT with either Shrinkage (mint_shr) or NOVELIST estimator (mint_n) return significant improvements over the base forecasts, especially in the 1-step-ahead horizon. However, the difference between their accuracy is hardly noticeable, with mint_n slightly outperform at the 1-step-ahead and mint_shr takes over in further forecasting horizons. mint_sample performs worse than base ARIMA models in short time series length, only to improve drastically in longer length. It is due to the sample covariance matrix gets more accurate as sample size gets larger.

These simulations

>>>>

In our second simulation study, we design a data‐generating process that contrasts “dense” and “sparse” correlation regimes among bottom‐level series, reflecting settings one might encounter in practice. Specifically, we construct two groups of fifty series each, with strong within-group dependencies throughout and either modest between‐group correlations (the dense scenario) or complete independence (the sparse scenario). These correlation matrices are depicted in Figure @fig-Sigma-50x2. Both scenarios share the same VAR(1) coefficient structure as in our previous simulations; only the innovation covariance changes. Such a setup mirrors real‐world contexts where, for example, sales within a product line may exhibit strong co‐movements, while those in a separate line operate nearly independently.

Rather than collapsing each group into a single aggregate as in previous practice, we impose a more intricate, multi‐level aggregation path to stress‐test reconciliation methods. We first form ten intermediate series by summing bottom‐level series in consecutive blocks of ten. These ten series are then assigned to three second‐level aggregates--groups of four, three, and four--and finally summed to produce a single top‐level series. This asymmetric hierarchy creates overlapping correlation patterns: some level‐2 series share bottom‐level groups, while others draw from both.
<!-- emulating practical scenarios such as regional sales aggregations that span multiple product categories or overlapping territories. -->

Figure @fig-sim-results-dsp presents out‐of‐sample mean squared error improvements over the base forecasts for each reconciliation strategy under both dense and sparse settings, with two panel lengths--54 and 304 observations--reserving the last four points for testing. In both short and long samples, MinT using either the shrinkage estimator (mint_shr) or the NOVELIST estimator (mint_n) delivers pronounced gains over incoherent ARIMA forecasts, particularly at the one‐step horizon where cross‐series correlations most directly inform the forecast adjustments. Although the two MinT variants perform almost indistinguishably overall, mint_n edges out mint_shr in the immediate horizon, whereas mint_shr slightly outperforms for longer horizons. By contrast, MinT with the raw sample covariance (mint_sample) suffers in small‐sample settings; as expected, its performance improves dramatically with 304 data points, since the sample covariance becomes more reliable with larger $n$. This highlights the practical necessity of regularized estimators in high‐dimensional, low‐sample contexts, a situation common in real applications where histories are short relative to the number of series.

We also experimented with alternative DGPs: varying group sizes, aggregation paths, and VAR(1) structure. However, to our surprise, most designs failed to distinguish mint_shr from mint_n in any substantive way. Their nearly identical performance under these synthetic scenarios suggests that our current simulation may not unveil the full advantages of the thresholding estimators. This finding motivates our turn to empirical data, where latent structural features and regime shifts, which we will discuss in the next section, may reveal performance differences.
